
---
title: "Project Assignment on Practical Machine Learning"
author: "Nayan Pandya"
date: "December 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project: Practical machine learning

### Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset)


## Data
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.


the content summary of weight lifting dataset:
reference from http://groupware.les.inf.puc-rio.br/har#weight_lifting_exercises 

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes. Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. We made sure that all participants could easily simulate the mistakes in a safe and controlled manner by using a relatively light dumbbell (1.25kg)."

## Read library to use packages for programme
```{r}

library(rpart)
library(caret)
library(ROSE)
library(Hmisc)
library(gridExtra)
library(GGally)
library(dummies)
library(car)
library(rpart.plot)
library(rattle)
library(randomForest)
library(knitr)

set.seed(3546)

```

# Read file from location

```{r}
setwd("C:/D drive/Digital strategy/field data/coursera/practical machinelearning/projectwork")
rm(list=ls())
dir()

testing <- read.csv("pml-testing.csv")
training <- read.csv("pml-training.csv")

```
# remove with NA columns more than 60%
```{r}

training <- training[,-which(colSums(is.na(training))>(0.6*length(training$classe)))]

```
# Partition of data into training and test set

```{r}

intrain <- createDataPartition(y=training$classe,p=0.75,list = FALSE)
train_data<- training[intrain,]
test_data <- training[-intrain,]


```
# Remove near zero variable from train and test data

```{r}
# checking near zero variable and remove it
nsv <- nearZeroVar(train_data,saveMetrics = TRUE)
train_data <- train_data[,-c(which(nsv$nzv== TRUE))]
rm(nsv)
nsv <- nearZeroVar(test_data,saveMetrics = TRUE)
test_data <- test_data[,-c(which(nsv$nzv== TRUE))]


``` 
# Convert all independent variable into numetic
```{r}
# convert all independent variable into numeric

train_num <- data.frame(lapply(train_data[,-59], function(x) as.numeric((x))))
train_data <- data.frame(train_num,train_data$classe)
test_num <- data.frame(lapply(test_data[,-59], function(x) as.numeric((x))))
test_data <- data.frame(test_num,test_data$classe)

```
# remove identification variable
``` {r}
train_data <- train_data[,-c(1:5)]
test_data <- test_data[,-c(1:5)]
# Final train data set
str(train_data)
# Final test data set 
str(test_data)
```

## check multicolineaty
```{r}

# Our response variable is catagorical variable so we should choose any independent variable as response variable and check multicolinearity

# Check VIF factor and remove variable which has VIF more than 5 but need to check iteratively
train_vif <- train_data[,-c(3,4,5,9,10,11,12,20,22,23,26,27,34,35,36,37,38,47,48,49)]

# Final dependent variables to use for prediction

str(train_vif)
dim(train_vif)
colnames(train_vif)
## convert test data with selected dependent variables
test_vif <- test_data
test_vif <- test_vif[,-c(3,4,5,9,10,11,12,20,22,23,26,27,34,35,36,37,38,47,48,49)]
dim(test_vif)
colnames(test_vif)
```
# Prediction with using Random Forest 

```{r}
library(randomForest)
set.seed(32343)
modelfit_rf <- randomForest(train_data.classe~.,data = train_vif,mtry=6,ntree=2001,importance=TRUE)
varImpPlot(modelfit_rf)

# Prediction with using random forest
confusionMatrix(predict(modelfit_rf,test_vif),test_vif$test_data.classe)

```
# Above graph is showing importance of factor for random forest



## prediction with using Grediant Boosting Method

```{r}
set.seed(32343)
modelfit_gbm <- train(train_data.classe~.,data=train_vif,method="gbm") 
# prediction with using GBM method
confusionMatrix(predict(modelfit_gbm,test_vif),test_vif$test_data.classe)

```

## Prediction with using classification method

```{r}
modelfit_rpart <- rpart(train_data.classe~.,data=train_vif,method = "class")
confusionMatrix(predict(modelfit_rpart, newdata=test_vif, type="class"),test_vif$test_data.classe)

```

## Comparison


Random forest Accuracy : 0.9973
Generalize boosting method Accuracy : 0.9888 
Classification method accuracy : 0.7359

Accuracy of random forest is high so final validation will be made using random forest.


## Following code is reducing testing column to length of training data set
```{r}

set.seed(32343)
testing <- testing[,-c(1:6)] 

testing <- testing[,-c(3:30)] 

testing <- testing[,-c(6:9)] 
testing <- testing[,-c(12:21)] 
testing <- testing[,-c(13,15,16,19:35)] 
testing <- testing[,-c(19:33)] 
testing <- testing[,-c(20:29)] 
testing <- testing[,-c(22:26)] 
testing <- testing[,-c(27:41,43:52)] 
testing <- testing[,-c(30:32)] 

colnames(testing)
colnames(test_vif)
table(testing$problem_id)
# Prediction with using random forest
confusionMatrix(predict(modelfit_rf,test_vif),test_vif$test_data.classe)

predict(modelfit_rf,testing)
```
